---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# MVMRLOVA

<!-- badges: start -->
<!-- badges: end -->

The goal of MVMRLOVA is to performs multivariable Mendelian randomization analysis based on latent phenotype of the outcome that explicitly exclude vertical pleiotropy effects via EM algorithm.
There is a single function:

-  **`MVMRLOVA()`**  main function that performs MVMR analysis and provides causal effect estimates.


## Installation

You can install the development version of **MVMRLOVA** from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("lamessad/MVMRLOVA")
```
To view the help pages for functions in this package, prepend the function name with a question mark.

```{r}
#library(MVMRLOVA)
#?MVMRLOVA()
```

## Example

Here is an example which shows how to apply MVMR-cML to infer the causal relationship from cholesterol levels (triglycerides, LDL and HDL) to CAD.

First extract GWAS summary data with \code{TwoSampleMR} package:

```{r}
#IVs extraction
library(TwoSampleMR)
exposuregwas_id = c(
  'ebi-a-GCST002216',
  'ebi-a-GCST002222')
outcomegwas_id = 'ebi-a-GCST005195'
exposure_dat = mv_extract_exposures(exposuregwas_id)
outcome_dat <- extract_outcome_data(exposure_dat$SNP,outcomegwas_id)
mvdat <- mv_harmonise_data(exposure_dat, outcome_dat)
betaX   <- mvdat$exposure_beta
colnames(betaX) <- c("TG", "LDL")
betaXse <- mvdat$exposure_se
betaY   <- as.matrix(mvdat$outcome_beta)  # make it a column matrix
betaYse <- as.matrix(mvdat$outcome_se)
ny=min(outcome_dat$samplesize.outcome)
#MVMRLOVA analysis
library(MVMRLOVA)
est <- MVMRLOVA(betaY, betaX, betaYse, betaXse, ny, gwas_p = 0.05, gwas_p2 = 5e-8, 100)
est$CausEst
```




